Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-09-19 paper/paper.tex
rewrote section describing movie pill plots Tested everything on quipu.

This simple cell shape is a good starting point for observing in
detail the dynamic interaction between the different stages of protein
that lead to their qualitatively distinct behavoir. Figure
\ref{image-p} shows the process in a series of frames taken from
simulation animations for both the deterministic and stochastic
models.  We have 'smeared out' the stochastic concentrations in a
manner meant to reproduce the images shown by diffraction limited
flourescence microscopy.  We do so using the two dimensional gaussian
approximation developed by Zhang \emph{et al.}~\cite{zhang2007gaussian}.  In
this approximation we use a numerical aperture value of 1.3, which is
the same as used by Mannik experimentally, and a wavelength of
$650nm$.  Each frame is 2.5 seconds ahead of the last, and each image
shows the concentration of a given state of protein (of the five
described in the reaction model) summed over the coordinate orthogonal
to the page.

Figure~\ref{image-p} begins about 300 seconds into the simulation and
shows one period of oscillation.  At $t=0$ the cytoplasmic MinE is
concentrated in the upper portion of the cell and is diffusing
downward, where it will react with and stick to wall-bound MinD that
is concentrated on the membrane in the lower portion of the cell.
This process is in fact already well on its way, as seen by the
significant concentration of wall bound MinD:MinE which is creeping
down to the lower corner, removing the MinD from the membrane as it
goes.  The removed MinD (cytoplasmic MinD:ADP) diffuses freely for a
time before spontenously undergoing nucleotide exchange and shifting
back into the MinD:ATP that is able to attach to the membrane.
Diffusing a small distance away from the bottom of the cell will bring
the protein into the MinE \fixme{collection ??}, where reactions will prevent it
from collecting for any significant time on the membrane.  Diffusing
further towards the upper end, however, while the MinE has not yet
collected there, will allow MinD:ATP to accumulate with the other
MinD:ATP, which is seen at seconds 10 through 15 in
Figure~\ref{image-p}.  Eventually the MinE creeps all the way down to
the bottom of the cell where it slowly runs out of wall bound MinD to
react with, before diffusing back up and starting again on the upper
portion of the cell. This is seen in seconds 15 through 20 in the
figure.  \fixme{There is another paragraph after this in the latex
  that I commented out because I thought we might be spending too much
  time on the description}

>>>>

Figure~\ref{image-p} shows a series of color plots of protein cycle
stage concentration for both the deterministic and stochastic
models. The cells shown are $4\micron$ in length, measured from end to
end.  We have 'smeared out' the stochastic concentrations in a manner
meant to reproduce the images shown by diffraction limited
flourescence microscopy.  We do so using the two dimensional gaussian
approximation developed by Zhang \emph{et
  al.}~\cite{zhang2007gaussian}.  In this approximation we use a
numerical aperture value of 1.3, which is the same as used by Mannik
experimentally, and a wavelength of $650nm$.  Each frame is 2.5
seconds ahead of the last, and each image shows the concentration of a
given state of protein (of the five described in the reaction model)
summed over the coordinate orthogonal to the page.

Figure~\ref{image-p} begins about 300 seconds into the simulation and
shows one period of oscillation.  At $t=0$ there is a high
concentration of MinD:ATP that has accumulated on the membrane in the
lower section of the cell. An important aspect of Huang's model is
that the MinD is attracted to and sticks to the membrane nonlinearly:
as it accumulates there it begins to 'recruit' other MinD that is
diffusing in the cytoplasm nearby, causing peaks in concentration to
build up on the walls.  Meanwhile, the MinE is creeping downward as it
reacts with the membrane-bound MinD, forming the MinD:MinE:ATP
complex, breaks it apart, and then diffuses downward a bit more before
it again reacts with membrane-bound MinD.  This process can seen by
the famous MinE 'rings' shown in the deterministic model plots, from 0
seconds to 4 seconds.  The appearence of these rings in the
deterministic model along side their less convincing appearence in the
stochastic model highlights an advantage of the deterministic model:
the idealization of deterministic data allows one to see macroscopic
patterns that might otherwise be missed.  In the stochastic model the
MinE still exhibits higher concentrations in the same regions
throughout the process, but what would ideally be a 'ring' pattern is
instead an assymetric collection of maxima.

During the formation of these rings, MinE that is diffusing in the
cytoplasm in the upper portion of the cell will naturally diffuse
downward, where there is membrane-bound MinD to react with, causing a
depletion of MinE in the upper portion of the cell.  As the MinE ring
converges upon the lower end of the cell, MinD that has been released
is able to diffuse upward, past the ring, while still in its MinD:ADP
state.  It then undergoes nucleotide exchange, resulting in MinD:ATP,
which is ready to accumulate on the walls in the upper portion of the
cell, where the MinE has been depleted.  This can be seen in seconds
10 through 20 in both models, and the subsequent MinE ring
formation and movement upward (beginning the same process in the
opposite direction) can be seen in seconds 15 through 23.

%% is concentrated in the
%% upper portion of the cell and is diffusing downward, where it will
%% react with and stick to wall-bound MinD that is concentrated on the
%% membrane in the lower portion of the cell.  This process is in fact
%% already well on its way, as seen by the significant concentration of
%% wall bound MinD:MinE which is creeping down to the lower corner,
%% removing the MinD from the membrane as it goes.  The removed MinD
%% (cytoplasmic MinD:ADP) diffuses freely for a time before spontenously
%% undergoing nucleotide exchange and shifting back into the MinD:ATP
%% that is able to attach to the membrane.  Diffusing a small distance
%% away from the bottom of the cell will bring the protein into the MinE
%% \fixme{collection ??}, where reactions will prevent it from collecting
%% for any significant time on the membrane.  Diffusing further towards
%% the upper end, however, while the MinE has not yet collected there,
%% will allow MinD:ATP to accumulate with the other MinD:ATP, which is
%% seen at seconds 10 through 15 in Figure~\ref{image-p}.  Eventually the
%% MinE creeps all the way down to the bottom of the cell where it slowly
%% runs out of wall bound MinD to react with, before diffusing back up
%% and starting again on the upper portion of the cell. This is seen in
%% seconds 15 through 20 in the figure.  \fixme{There is another
%%   paragraph after this in the latex that I commented out because I
%%   thought we might be spending too much time on the description}
